Effective multi-body SU($N$)-symmetric interactions of ultracold fermionic atoms on a 3-D lattice

### 2019-02-18 multi_body_interactions/papers/theory/reviews_NJP/revision_summary.tex
final versions of responses and revisions to NJP

\setcounter{point}{1}
\begin{enumerate}[label=(R2.\arabic{point}.\arabic*)]
\item
\end{enumerate}

@@

  Both examples are ``correct'' due to the fact that
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is self-adjoint, but the latter
  ordering convention is used throughout the rest of the manuscript:
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is gives the coupling constant
  for destroying two atoms in state $\p{\mu q}+\p{\nu r}$, and
  creating them in $\p{\rho s}+\p{\sigma t}$.

@@

@@

>>>>

\setcounter{point}{1}
\begin{enumerate}[label=(R2.\arabic{point}.\arabic*)]
  \step
\item We have made the caption of figure 1 more self-contained,
  replacing the text (page 11, line 13):

  \red{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake
    presentation, ...}

  \green{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  Diagrams are
    read from left to right to construct a sequence of operators from
    right to left.  Solid (dashed) lines represent field operators
    acting on the lowest (arbitrary) motional states.  For the sake
    presentation, ...}

  \step
\item We have rewritten our discussion of figures 4 and 5, replacing
  the text (page 23, line 54):

  \red{Identifying peaks in excitation spectra such as in figure 3
    constitutes a measurement of many-body excitation energies, which
    was performed in ref.~[34] to detect signatures of effective
    multi-body interactions.  Figure 4 shows a comparison between
    experimental measurements of the many-body excitation energies
    $\Delta_{NX}$ for $N\in\set{3,4,5}$ at various lattice depths
    $\U$, and the corresponding values of $\Delta_{NX}$ predicted by
    the low-energy effective theory at different orders in the
    coupling constants.  A summary of the data in figure 4 is provided
    in figure 5.  These figures show a clear improvement of the
    low-energy effective theory we have developed with increasing
    order in the coupling constants, in particular demonstrating the
    need to consider multi-body interactions in a single-band model of
    the atoms without motional excitations.}

  \red{We expect that the dominant source of error in the third-order
    effective theory comes from the fact that we neglect the
    inter-site matrix elements of all Hamiltonians.  This source of
    error is discussed in Appendix E, and leads to theoretical
    uncertainties represented by error bars in figure 4.  As these
    uncertainties are sometimes unable to account for the residual
    disagreement between experimental measurements and the effective
    theory, better agreement requires either a better understanding of
    experimental procedures and error, a better understanding of
    theoretical uncertainties, or consideration of yet higher-order
    terms in the effective theory.  Due to a combinatorial explosion
    in the number of diagrams which appear at increasing perturbative
    orders of the effective theory, we need more systematic methods to
    compute effective multi-body Hamiltonians at fourth order; we
    therefore leave this calculation for future work.}

  \green{Identifying peaks in excitation spectra such as in figure 3
    constitutes a measurement of many-body excitation energies, which
    was performed in ref.~[34] to detect signatures of effective
    multi-body interactions.  Figure 4 shows a comparison between (i)
    experimental measurements of the many-body excitation energies
    $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
    various mean lattice depths $\U$, and (ii) the corresponding
    values of $\Delta_{NX}$ predicted by the low-energy effective
    theory at different orders in the coupling constants.  A known
    source of error in our effective theory comes from neglecting the
    inter-site matrix elements of all Hamiltonians.  This error is
    discussed in Appendix E, and leads to theoretical uncertainties
    represented by error bars on the $\O\p{G^3}$ theory in figure 4.
    A summary of figure 4 is provided in figure 5.  We note that
    many-body interaction energy shifts are smaller for asymmetric
    ($-$) states than symmetric ($+$) ones due to the competition
    between contributions of opposite sign in the asymmetric case (see
    rows 2 and 3 of table I, where as a consequence of positive
    scattering lengths in the case of $^{87}$Sr, all $U_{MX}$ for
    fixed $M$ have the same sign).  This competition is a many-body
    analogue of the two-body case with a competition between direct
    and exchange terms in the interaction energies of singly-excited
    states.}

  \green{The results in figures 4 and 5 highlight a few important
    points about ultracold, high-density ${}^{87}$Sr experiments and
    our low-energy effective theory.  First, these experiments exhibit
    clear signatures of multi-body interactions, as evidenced by a
    stark disagreement between the observed many-body excitation
    energies $\Delta_{NX}$ and those that are predicted by the
    two-body $\O\p{G}$ theory.  Multi-body interactions are thus
    crucial for understanding these high-density experiments in the
    context of a single-band Hubbard model, which naturally arises in
    the zero-temperature limit when all atoms occupy their motional
    ground state.  Second, the inter-atomic interactions in these
    experiments are strong enough to require going beyond leading
    order for the description of multi-body interactions in the
    low-energy effective theory.  The formally small quantities
    organizing our perturbation theory are two-body ground-state
    interaction energies (proportional to the couplings $G_X$) divided
    by the spectral gap of the single-atom Hamiltonian $H_0$.  These
    reduced (dimensionless) interaction energies vary from
    $\sim0.05-0.15$ in the parameter regimes of the ${}^{87}$Sr
    experiments considered here (see Appendix F).  As experiments
    begin to operate at higher atom densities with amplified
    interaction effects, reliably predicting interaction energies may
    require going to yet higher orders in perturbation theory.  Due to
    a combinatorial explosion of the number of diagrams which appear
    at increasing orders in the effective theory, however, we need
    more systematic methods to compute effective multi-body
    Hamiltonians at fourth order.  In any case, we are agnostic as to
    whether such a calculation would provide better agreement between
    experiment and theory without first performing a detailed analysis
    of systematic errors.}

  Note that the Appendix F referenced above is not the same as that in
  the original version of our manuscript, but a new appendix that we
  have added in the revised version.

  \step
\item We have added the following appendix to our manuscript:

  \setcounter{figure}{0}
  \renewcommand\thefigure{F.\arabic{figure}}

  \begin{figure}
    \centering
    \includegraphics{../figures/perturbative_params.pdf}
    \caption{\footnotesize Dependence of the perturbative parameters
      $KG_X/\Delta$ on the lattice depth $\U$.}
  \end{figure}

  \green{{\bf Appendix F: Perturbative parameters for the effective
      theory}}

  \green{The perturbative effective theory developed in Section III is
    organized in powers of the coupling constants $G_X$.  The formally
    small, dimensionless quantities for this perturbation theory are
    the two-body interaction energies $K G_X$ divided by the spectral
    gap $\Delta$ of the non-interacting Hamiltonian $H_0$.  Here $K$
    is a ground-state two-body overlap integral and $G_X$ is a
    coupling constant.  The quantities $K$, $G_X$, and $\Delta$ all
    depend on the lattice depth $\U$.  Figure F.1 shows these
    parameters for the case of ${}^{87}$Sr atoms with
    $X\in\set{\g\g, \e\g_-, \e\g_+, \e\e}$ at lattice depths
    $\U\in\sp{30,80}E_{\t{R}}$, where
    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
    recoil energy of the atoms.  The fact that these perturbative
    parameters grow with increasing lattice depth $\U$ is a
    consequence of the fact that the overlap integral $K$ grows faster
    with $\U$ than the spectral gap $\Delta$.  In the case of a
    harmonic trap with angular frequency $\omega$, for example, by
    dimensional analysis these parameters would be
    \begin{align*}
      \f{K_{\t{HO}} G_X}{\omega}
      = \f{G_X}{\omega} \int \d^3 x~\abs{\phi_{\t{HO}}}^4
      = \f{G_X}{\omega} \sp{\int \d x~
        \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3
      \sim \omega^{1/2},
      \tag{F1}
    \end{align*}
    where we assumed that the coupling constants $G_X$ vary weakly
    with $\omega$.  While this result may seem to suggest that the
    low-energy effective theory should become better at smaller
    lattice depths, smaller lattice depths also result in increased
    theoretical uncertainties from the growing relevance of the
    inter-site effects discussed in Appendix E.}

@@

  Both examples are ``correct'' due to the fact that
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is self-adjoint, but the latter
  ordering convention is used throughout the rest of the manuscript:
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ gives the coupling constant for
  destroying two atoms in state $\p{\mu q}+\p{\nu r}$, and creating
  them in $\p{\rho s}+\p{\sigma t}$.

@@

\item In table I, we have change the following sentence of the caption
  (page 20, line 7):

  \red{The corresponding $N$-body eigenvalues $E_{NX}^{(M)}$ of $H_M$
    are given in terms of the coefficients $U_{MX}$ as appearing in
    (42), in addition to the $M$-body eigenvalues $E_{MX}^{(M)}$.}

  \green{The corresponding $N$-body eigenvalues $E_{NX}^{(M)}$ of
    $H_M$ are given in terms of the coefficients $U_{MX}$ as appearing
    in (44) (first three rows), in addition to the $M$-body
    eigenvalues $E_{MX}^{(M)}$ (last three rows).}

  Note that the equation referenced in this text has only changed due
  to the addition of two more equations in the revised version of our
  manuscript.  In addition, we have made explicit that this table
  applies for $M\le N$, by changing the column header (page 20, line
  13) ``\red{$H_M$ eigenvalue}'' to ``\green{$H_M$ eigenvalue
    ($M\le N$)}''.

@@

\item We have fixed an error in our list of low-excitation $M$-body
  Hamiltonian coefficients, replacing
  \red{$\p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}$} by
  \green{$\p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}$} in Eqs.~(F8),
  (F9), and (F10) on page 41.  Note that these equations are
  respectively (G8), (G9), and (G10) in the revised manuscript.
